---
title: "Monetary Policy Example"
author: "Michael Wieck-Sosa"
date: "12/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Load packages**
```{r}
#installation of the BMR package is a bit of a pain
# 1. install.packages("devtools")
# 2. library(devtools)
# 3. #Sys.setenv("R_REMOTES_NO_ERRORS_FROM_WARNINGS" = "true")
# 4. install_github("kthohr/BMR", INSTALL_opts="--no-multiarch")
# May need to also install Rtools if on Windows https://cran.r-project.org/bin/windows/Rtools/
# and for Mac users https://cran.r-project.org/bin/macosx/tools/
library(BMR)

# install.packages("fractal")
library(fractal)

library(reshape2)
library(dplyr)
library(zoo)
```

# Minnesota Prior Example
```{r}
# Get macroeconomic monthly data from January 1959 to July 2020

# from https://fred.stlouisfed.org/series/UNRATE
unrate <- read.csv("https://fred.stlouisfed.org/graph/fredgraph.csv?bgcolor=%23e1e9f0&chart_type=line&drp=0&fo=open%20sans&graph_bgcolor=%23ffffff&height=450&mode=fred&recession_bars=on&txtcolor=%23444444&ts=12&tts=12&width=748&nt=0&thu=0&trc=0&show_legend=yes&show_axis_titles=yes&show_tooltip=yes&id=UNRATE&scale=left&cosd=1959-01-01&coed=2020-07-01&line_color=%234572a7&link_values=false&line_style=solid&mark_type=none&mw=3&lw=2&ost=-99999&oet=99999&mma=0&fml=a&fq=Monthly&fam=avg&fgst=lin&fgsnd=2020-02-01&line_index=1&transformation=lin&vintage_date=2020-12-11&revision_date=2020-12-11&nd=1948-01-01")
# no transformation

# from https://fred.stlouisfed.org/series/PCEPILFE
pcepilfe <- read.csv("https://fred.stlouisfed.org/graph/fredgraph.csv?bgcolor=%23e1e9f0&chart_type=line&drp=0&fo=open%20sans&graph_bgcolor=%23ffffff&height=450&mode=fred&recession_bars=on&txtcolor=%23444444&ts=12&tts=12&width=748&nt=0&thu=0&trc=0&show_legend=yes&show_axis_titles=yes&show_tooltip=yes&id=PCEPILFE&scale=left&cosd=1959-01-01&coed=2020-07-01&line_color=%234572a7&link_values=false&line_style=solid&mark_type=none&mw=3&lw=2&ost=-99999&oet=99999&mma=0&fml=a&fq=Monthly&fam=avg&fgst=lin&fgsnd=2020-02-01&line_index=1&transformation=lin&vintage_date=2020-12-11&revision_date=2020-12-11&nd=1959-01-01")
# no transformation

# from https://fred.stlouisfed.org/series/PCEPI
pcepi <- read.csv("https://fred.stlouisfed.org/graph/fredgraph.csv?bgcolor=%23e1e9f0&chart_type=line&drp=0&fo=open%20sans&graph_bgcolor=%23ffffff&height=450&mode=fred&recession_bars=on&txtcolor=%23444444&ts=12&tts=12&width=748&nt=0&thu=0&trc=0&show_legend=yes&show_axis_titles=yes&show_tooltip=yes&id=PCEPI&scale=left&cosd=1959-01-01&coed=2020-07-01&line_color=%234572a7&link_values=false&line_style=solid&mark_type=none&mw=3&lw=2&ost=-99999&oet=99999&mma=0&fml=a&fq=Monthly&fam=avg&fgst=lin&fgsnd=2020-02-01&line_index=1&transformation=lin&vintage_date=2020-12-11&revision_date=2020-12-11&nd=1959-01-01")
# no transformation

# from https://fred.stlouisfed.org/series/GDPC1#0
real_gdp <- read.csv("https://fred.stlouisfed.org/graph/fredgraph.csv?bgcolor=%23e1e9f0&chart_type=line&drp=0&fo=open%20sans&graph_bgcolor=%23ffffff&height=450&mode=fred&recession_bars=on&txtcolor=%23444444&ts=12&tts=12&width=748&nt=0&thu=0&trc=0&show_legend=yes&show_axis_titles=yes&show_tooltip=yes&id=GDPC1&scale=left&cosd=1959-01-01&coed=2020-07-01&line_color=%234572a7&link_values=false&line_style=solid&mark_type=none&mw=3&lw=2&ost=-99999&oet=99999&mma=0&fml=a&fq=Quarterly&fam=avg&fgst=lin&fgsnd=2020-02-01&line_index=1&transformation=lin&vintage_date=2020-12-11&revision_date=2020-12-11&nd=1947-01-01")
real_gdp$DATE <- as.Date(real_gdp$DATE)
# use a spline method to convert quarterly real_gdp to monthly real_gdp
months <- lapply(X = real_gdp$DATE, FUN = seq.Date, by = "month", length.out = 3)
months <- data.frame(DATE = do.call(what = c, months))
monthly_data <- left_join(x = months, y = real_gdp, by = "DATE")
monthly_data$GDPC1 <- na.spline(object = monthly_data$GDPC1)
real_gdp <- monthly_data
# log of GDP
real_gdp$GDPC1 <- log(real_gdp$GDPC1)
# get rid of last 2 rows from spline
real_gdp <- real_gdp[-c(740,741),]

# combine macroeconomic variables
df <- data.frame(cbind(real_gdp,unrate$UNRATE,pcepilfe$PCEPILFE,pcepi$PCEPI))
names(df) <- c("DATE","REAL_GDP","UNRATE","PCEPILFE","PCEPI")

# from https://www.cbo.gov/data/budget-economic-data#6
# the very bottom left of the page for the most up to date 
# potential GDP forecasts by the Congressional Budget Office
real_potential_gdp <- read.csv("C:/Users/mykul/OneDrive/Documents/stat429/proj/Quarterly_July20.csv")
real_potential_gdp <- real_potential_gdp[,c("date","real_potential_gdp")]
real_potential_gdp$date <- seq(as.Date("1949-01-01"), as.Date("2030/10/01"), by = "quarter")
# use spline method to get monthly potential GDP
months <- lapply(X = real_potential_gdp$date, FUN = seq.Date, by = "month", length.out = 3)
months <- data.frame(date = do.call(what = c, months))
monthly_data <- left_join(x = months, y = real_potential_gdp, by = "date")
monthly_data$real_potential_gdp <- na.spline(object = monthly_data$real_potential_gdp)
real_potential_gdp <- monthly_data
# log of potential GDP
real_potential_gdp$real_potential_gdp <- log(real_potential_gdp$real_potential_gdp)

```

```{r}
# plot time-series 
gtsplot(df[, 2:5], dates = df[, 1])

```


```{r}
# n.taper is the number of sinusoidal tapers to develop eigenspectra for each block of the time series because 5 is the minimum value
# n.block is the number of blocks to divide the time series, ideally should be divisible by n.tapers so to use all the data, so I chose 10 
stationarity(as.matrix(df[,2:5]),n.taper = 5,n.block = 10)
# reject Priestley-Subba Rao stationarity Test null hypothesis that time-series is stationary at alpha = 0.05 significance level

```


```{r}
# create new R object of BVAR with Minnesota Prior
myBvar <- new(bvarm)

# build BVAR with Minnesota prior, with constant, and with four lags
myBvar$build(data.matrix(df[,2:5]),TRUE,4)

# Citations for choosing BVAR model setup follows 
# the set-up of the Economics academic paper Canova (2007)
# set priors as the same for each variable to get random-walk for 
# the variables which are non-stationary (which I tested above)
coefficientPrior <- c(0.9,0.9,0.9,0.9)

myBvar$prior(coefficientPrior, # prior mean value
               1,    # var type
               1,    # decay type
               0.2,  # hyperparameter 1
               0.5,  # hyperparameter 2
               10^5, # hyperparameter 3
               1.0)  # hyperparameter 4

# 10,000 draws for Gibbs sampler 
myBvar$gibbs(10000)


```

```{r}
# density functions for the parameter estimates
plot(myBvar, var_names = colnames(df)[-1], save = FALSE)
```


```{r}
# impulse response function
IRF.Rcpp_bvars(myBvar,20,var_names=colnames(df)[-1],save=FALSE)
```


```{r}
forecastEstimates <- forecast(myBvar,shocks=TRUE,var_names=colnames(df)[-1],back_data=10,save=FALSE)

```


```{r}
# get forecast estimates into a dataframe
df_forecasted <- data.frame(forecastEstimates$forecast_mean)
names(df_forecasted) <- c("REAL_GDP","UNRATE","PCEPILFE","PCEPI")

df_forecasted$DATE <- seq(as.Date("2020-08-01"),as.Date("2022-03-01"),by="month")
df_forecasted <- df_forecasted %>% select("DATE", everything())


```


```{r}
# add real potential gdp as a column to both series 
# original series: January 1, 1959 to July 1, 2020
df$real_potential_gdp <- real_potential_gdp[121:859,2]

# forecast series: August 1, 2020 to March 1, 2022
df_forecasted$real_potential_gdp <- real_potential_gdp[860:879,2]

# combine both series 
df_fullSeries <- rbind(df,df_forecasted)

```


```{r}
# calculate year-by-year PCE inflation rate
df_fullSeries$PCE_INFLATION_RATE <- (df_fullSeries$PCEPILFE - lag(df_fullSeries$PCEPILFE,12)) / lag(df_fullSeries$PCEPILFE,12)

# remove first 12 rows
df_fullSeries <- df_fullSeries[-c(1:12),]

```

```{r}
# calculate percent deviation of real GDP from real potential GDP target
df_fullSeries$percentDeviation_REALGDP <- (df_fullSeries$real_potential_gdp - df_fullSeries$REAL_GDP) / df_fullSeries$REAL_GDP
0
```


```{r}
# Taylor Rule
# see https://www.brookings.edu/blog/ben-bernanke/2015/04/28/the-taylor-rule-a-benchmark-for-monetary-policy/
# r = p + .5y + .5(p – 2) + 2
# where
# r = the federal funds rate
# p = the rate of inflation
# y = the percent deviation of real GDP from a target

df_fullSeries$fedFundsRate <- df_fullSeries$PCE_INFLATION_RATE + 0.5*df_fullSeries$percentDeviation_REALGDP + 0.5*(df_fullSeries$PCE_INFLATION_RATE - 2) + 2

tail(df_fullSeries[c("DATE","fedFundsRate")],20)

```


# Normal-Inverse Wishart Prior Example

```{r}
# Get macroeconomic monthly data from January 1959 to July 2020

# from https://fred.stlouisfed.org/series/UNRATE
unrate <- read.csv("https://fred.stlouisfed.org/graph/fredgraph.csv?bgcolor=%23e1e9f0&chart_type=line&drp=0&fo=open%20sans&graph_bgcolor=%23ffffff&height=450&mode=fred&recession_bars=on&txtcolor=%23444444&ts=12&tts=12&width=748&nt=0&thu=0&trc=0&show_legend=yes&show_axis_titles=yes&show_tooltip=yes&id=UNRATE&scale=left&cosd=1959-01-01&coed=2020-07-01&line_color=%234572a7&link_values=false&line_style=solid&mark_type=none&mw=3&lw=2&ost=-99999&oet=99999&mma=0&fml=a&fq=Monthly&fam=avg&fgst=lin&fgsnd=2020-02-01&line_index=1&transformation=lin&vintage_date=2020-12-11&revision_date=2020-12-11&nd=1948-01-01")
# no transformation

# from https://fred.stlouisfed.org/series/PCEPILFE
pcepilfe <- read.csv("https://fred.stlouisfed.org/graph/fredgraph.csv?bgcolor=%23e1e9f0&chart_type=line&drp=0&fo=open%20sans&graph_bgcolor=%23ffffff&height=450&mode=fred&recession_bars=on&txtcolor=%23444444&ts=12&tts=12&width=748&nt=0&thu=0&trc=0&show_legend=yes&show_axis_titles=yes&show_tooltip=yes&id=PCEPILFE&scale=left&cosd=1959-01-01&coed=2020-07-01&line_color=%234572a7&link_values=false&line_style=solid&mark_type=none&mw=3&lw=2&ost=-99999&oet=99999&mma=0&fml=a&fq=Monthly&fam=avg&fgst=lin&fgsnd=2020-02-01&line_index=1&transformation=lin&vintage_date=2020-12-11&revision_date=2020-12-11&nd=1959-01-01")
# no transformation

# from https://fred.stlouisfed.org/series/PCEPI
pcepi <- read.csv("https://fred.stlouisfed.org/graph/fredgraph.csv?bgcolor=%23e1e9f0&chart_type=line&drp=0&fo=open%20sans&graph_bgcolor=%23ffffff&height=450&mode=fred&recession_bars=on&txtcolor=%23444444&ts=12&tts=12&width=748&nt=0&thu=0&trc=0&show_legend=yes&show_axis_titles=yes&show_tooltip=yes&id=PCEPI&scale=left&cosd=1959-01-01&coed=2020-07-01&line_color=%234572a7&link_values=false&line_style=solid&mark_type=none&mw=3&lw=2&ost=-99999&oet=99999&mma=0&fml=a&fq=Monthly&fam=avg&fgst=lin&fgsnd=2020-02-01&line_index=1&transformation=lin&vintage_date=2020-12-11&revision_date=2020-12-11&nd=1959-01-01")
# no transformation

# from https://fred.stlouisfed.org/series/GDPC1#0
real_gdp <- read.csv("https://fred.stlouisfed.org/graph/fredgraph.csv?bgcolor=%23e1e9f0&chart_type=line&drp=0&fo=open%20sans&graph_bgcolor=%23ffffff&height=450&mode=fred&recession_bars=on&txtcolor=%23444444&ts=12&tts=12&width=748&nt=0&thu=0&trc=0&show_legend=yes&show_axis_titles=yes&show_tooltip=yes&id=GDPC1&scale=left&cosd=1959-01-01&coed=2020-07-01&line_color=%234572a7&link_values=false&line_style=solid&mark_type=none&mw=3&lw=2&ost=-99999&oet=99999&mma=0&fml=a&fq=Quarterly&fam=avg&fgst=lin&fgsnd=2020-02-01&line_index=1&transformation=lin&vintage_date=2020-12-11&revision_date=2020-12-11&nd=1947-01-01")
real_gdp$DATE <- as.Date(real_gdp$DATE)
# use a spline method to convert quarterly real_gdp to monthly real_gdp
months <- lapply(X = real_gdp$DATE, FUN = seq.Date, by = "month", length.out = 3)
months <- data.frame(DATE = do.call(what = c, months))
monthly_data <- left_join(x = months, y = real_gdp, by = "DATE")
monthly_data$GDPC1 <- na.spline(object = monthly_data$GDPC1)
real_gdp <- monthly_data
# log of GDP
real_gdp$GDPC1 <- log(real_gdp$GDPC1)
# get rid of last 2 rows from spline
real_gdp <- real_gdp[-c(740,741),]

# combine macroeconomic variables
df <- data.frame(cbind(real_gdp,unrate$UNRATE,pcepilfe$PCEPILFE,pcepi$PCEPI))
names(df) <- c("DATE","REAL_GDP","UNRATE","PCEPILFE","PCEPI")

# from https://www.cbo.gov/data/budget-economic-data#6
# the very bottom left of the page for the most up to date 
# potential GDP forecasts by the Congressional Budget Office
real_potential_gdp <- read.csv("C:/Users/mykul/OneDrive/Documents/stat429/proj/Quarterly_July20.csv")
real_potential_gdp <- real_potential_gdp[,c("date","real_potential_gdp")]
real_potential_gdp$date <- seq(as.Date("1949-01-01"), as.Date("2030/10/01"), by = "quarter")
# use spline method to get monthly potential GDP
months <- lapply(X = real_potential_gdp$date, FUN = seq.Date, by = "month", length.out = 3)
months <- data.frame(date = do.call(what = c, months))
monthly_data <- left_join(x = months, y = real_potential_gdp, by = "date")
monthly_data$real_potential_gdp <- na.spline(object = monthly_data$real_potential_gdp)
real_potential_gdp <- monthly_data
# log of potential GDP
real_potential_gdp$real_potential_gdp <- log(real_potential_gdp$real_potential_gdp)

```

```{r}
# plot time-series 
gtsplot(df[, 2:5], dates = df[, 1])

```


```{r}
# n.taper is the number of sinusoidal tapers to develop eigenspectra for each block of the time series because 5 is the minimum value
# n.block is the number of blocks to divide the time series, ideally should be divisible by n.tapers so to use all the data, so I chose 10 
stationarity(as.matrix(df[,2:5]),n.taper = 5,n.block = 10)
# reject Priestley-Subba Rao stationarity Test null hypothesis that time-series is stationary at alpha = 0.05 significance level

```


```{r}
# create new R object of BVAR with normal-inverse-Wishart Prior
myBvar <- new(bvarinw)

# build BVAR with normal-inverse Wishart prior, with constant and 4 lags
coefficientPrior <- c(0.9,0.9,0.9,0.9)
xBetaValue <- 4
xSigmaValue <- 1
gammaValue <- 4
numLags <- 4
addConstant <- TRUE

myBvar$build(data.matrix(df[,2:5]),addConstant,numLags)

# Citations for choosing BVAR model setup follows 
# the set-up of the Economics academic paper Canova (2007)
# set priors as the same for each variable to get random-walk for 
# the variables which are non-stationary (which I tested above)
myBvar$prior(coefficientPrior,xBetaValue,xSigmaValue,gammaValue)
bvar_obj$prior(coef_prior,XiBeta,XiSigma,gamma)

# 10,000 draws for Gibbs sampler 
myBvar$gibbs(10000,5000)



```

```{r}
# density functions for the parameter estimates
plot(myBvar, var_names = colnames(df)[-1], save = FALSE)
```


```{r}
# impulse response function
IRF.Rcpp_bvars(myBvar,3,var_names=colnames(df)[-1],save=FALSE)
```


```{r}
forecastEstimates <- forecast(myBvar,3,shocks=TRUE,var_names=colnames(df)[-1],back_data=10,save=FALSE)

```


```{r}
# get forecast estimates into a dataframe
df_forecasted <- data.frame(forecastEstimates$forecast_mean)
names(df_forecasted) <- c("REAL_GDP","UNRATE","PCEPILFE","PCEPI")

df_forecasted$DATE <- seq(as.Date("2020-08-01"),as.Date("2022-03-01"),by="month")
df_forecasted <- df_forecasted %>% select("DATE", everything())


```


```{r}
# add real potential gdp as a column to both series 
# original series: January 1, 1959 to July 1, 2020
df$real_potential_gdp <- real_potential_gdp[121:859,2]

# forecast series: August 1, 2020 to March 1, 2022
df_forecasted$real_potential_gdp <- real_potential_gdp[860:879,2]

# combine both series 
df_fullSeries <- rbind(df,df_forecasted)

```


```{r}
# calculate year-by-year PCE inflation rate
df_fullSeries$PCE_INFLATION_RATE <- (df_fullSeries$PCEPILFE - lag(df_fullSeries$PCEPILFE,12)) / lag(df_fullSeries$PCEPILFE,12)

# remove first 12 rows
df_fullSeries <- df_fullSeries[-c(1:12),]

```

```{r}
# calculate percent deviation of real GDP from real potential GDP target
df_fullSeries$percentDeviation_REALGDP <- (df_fullSeries$real_potential_gdp - df_fullSeries$REAL_GDP) / df_fullSeries$REAL_GDP
```


```{r}
# Taylor Rule
# see https://www.brookings.edu/blog/ben-bernanke/2015/04/28/the-taylor-rule-a-benchmark-for-monetary-policy/
# r = p + .5y + .5(p – 2) + 2
# where
# r = the federal funds rate
# p = the rate of inflation
# y = the percent deviation of real GDP from a target

df_fullSeries$fedFundsRate <- df_fullSeries$PCE_INFLATION_RATE + 0.5*df_fullSeries$percentDeviation_REALGDP + 0.5*(df_fullSeries$PCE_INFLATION_RATE - 2) + 2

tail(df_fullSeries[c("DATE","fedFundsRate")],20)

View(df_fullSeries)

```



**Citations:**
1. https://www.kthohr.com/bmr/BMR.pdf 
2. https://kevinkotze.github.io/ts-9-tut/
3. https://www.r-econometrics.com/timeseries/bvar/
4. https://rdrr.io/github/kthohr/BMR/man/bvarw.html
5. https://rdrr.io/github/kthohr/BMR/man/bvarw.html
6. https://research.stlouisfed.org/econ/mccracken/fred-databases/
7. https://fred.stlouisfed.org/
8. https://www.brookings.edu/blog/ben-bernanke/2015/04/28/the-taylor-rule-a-benchmark-for-monetary-policy/

**Some comments:**
1. The BMR package has changed since these guides were written and therefore none of them ran successfully so I had to tinker around with R for a while to use the new functions that the package supports with respect to R C++ objects. So I had to do something different to get the BMR package examples to work at all.
2. I changed the parameter values for the model setup a bit since we are using different variables, but stayed in the same general direction.
3. I used more different variables (GDP, Unemployment Rate, PCE Inflation) from FRED to get the most recent data, and I used a spline to interpolate the quarterly GDP into monthly GDP. Also, we get the potential GDP for the next 10 years from the congressional budget office. Lastly, we calculated the projected federal funds rate using the Taylor Rule.
4. I followed the Macroeconomic theory papers for determining the BVAR models.
- Canova, Fabio. "G-7 inflation forecasts: Random walk, Phillips curve or what else?." Macroeconomic Dynamics 11.1 (2007): 1-30 for their proposed setup of the BVAR model. 
- Canova, Fabio, Methods for Applied Macroeconomic Research, Princeton, New Jersey: Princeton University Press, 2007.
- Koop, Gary and Dimitris Korobilis, “Bayesian Multivariate Time Series Methods for Empirical Macroeconomics,” Mimeo, 2010.
5. I added a section on the Normal-inverse-Wishart Prior which no example had implemented.
























